The soundness proof proves that if a rule is successfully derived using the
LLAM, then it can also be derived in HLD. Since HLD semantics are so close to
linear logic, we prove that LM has a determined, correct, proof search behavior
when executing programs. However, the completeness theorem cannot be proven
since LLAM lacks the non-determinism present in HLD.

\paragraph{Initial Definitions} We first define equality between two multi-sets
of terms. Two multi-sets $A$ and $B$ are equal, $\feq{A}{B}$, when they have the
same constituent atoms\footnote{Also extends to atoms with variables, in which
there is a valid substitution to make them equal.}. This definition will be
useful to define what it means for a frame and state to be well-formed.

{\scriptsize
\[
\infer[\equiv p]{\feq{p, A}{q, B}}{\feq{A\theta_1}{B\theta_2} &
p\theta_1 \triangleq q\theta_2}
\tab
\infer[\equiv \bang p]{\feq{\bang p, A}{\bang q, B}}{\feq{A\theta_1}{B\theta_2}
& \bang p \theta_1 \triangleq \bang q \theta_2}
\tab
\infer[\equiv \one~L]{\feq{\one, A}{B}}{\feq{A}{B}}
\tab
\infer[\equiv \one~R]{\feq{A}{\one, B}}{\feq{A}{B}}
\]

\[
\infer[\equiv \cdot]{\feq{\cdot}{\cdot}}{}
\tab
\infer[\equiv \otimes~L]{\feq{A \otimes B, C}{D}}{\feq{A, B, C}{D}}
\tab
\infer[\equiv \otimes~R]{\feq{A}{B \otimes C, D}}{\feq{A}{B, C, D}}
\]
\vspace{-5mm}
}

\iffalse
\begin{theorem}[Match equivalence]
If two multi-sets are equivalent, $\feq{A_1, \dotsc, A_n}{B_1, \dotsc, B_m}$,
  and we can match $A_1 \otimes \dotsb \otimes A_n$ in HLD such that
  $\mz{\Gamma}{\Delta}{A_1 \otimes \dotsb \otimes A_n}$ then
     $\mz{\Gamma}{\Delta}{B_1 \otimes \dotsb \otimes B_m}$ is also true.
\end{theorem}
\begin{proof}
By straightforward induction on the first assumption.
\end{proof}
\fi

\begin{definition}[Well-formed frame]
  Consider a triplet $A; \Gamma; \Delta_{N}$ where $A$ is a term, $\Gamma$ is a
  set of persistent resources and $\Delta_{N}$ a multi-set of linear
  resources. A linear frame $f = \lframe{\Delta,
        p_1}{\Delta''}{p}{\Omega}{\Delta'}{\Omega'}$ is well-formed iff:

\begin{enumerate}
  \item $\feq{p, \Omega, \Omega'}{A}$ (the remaining terms and 
           matched terms are equivalent to body $A$);
  \item $\Delta' = \Delta'_1, \dotsc, \Delta'_n$ and $\Omega' =
  \Omega'_1 \otimes \dotsb \otimes \Omega'_n$
  \item $\Delta, \Delta'', \Delta, p_1 = \Delta_{N}$ (available facts, candidate
        facts for $p$, consumed facts and the linear fact used for $p$,
        respectively, are the same as the initial $\Delta_{N}$);
  \item $\mz{\Gamma}{\Delta'}{\Omega'}$ is valid.
\end{enumerate}
\end{definition}

The well-formedness of a continuation stack follows the above definition for all
frames. Moreover, each adjacent continuation frame is related in the sense that
$\Omega'$ of the first frame is always smaller than the $\Omega'$ of the next frame.
Identical relations apply to $\Delta'$ (increases) and $\Omega$ (decreases).

We also need to define what it means for a LLAM state to be well-formed. For
matching states in particular, we need to ensure that the proof term
$\mz{\Gamma}{\Delta}{\Omega}$ is related to the term being matched. We show
the body match definition below.

\begin{definition}[Well-formed body match]
  $\matstate{A \lolli B}{\rulestk}{\lstack{C}}{\Gamma}{\Delta}{\Omega}{\Delta'
  \rightarrow \Omega'}$ is well-formed in relation to a triplet $A; \Gamma;
  \Delta_{N}$ iff:

  \begin{itemize}
     \item $\Delta, \Delta' = \Delta_{N}$ and $\feq{\Omega', \Omega}{A}$ and
        $\mz{\Gamma}{\Delta'}{\Omega'}$ is valid;
     \item $\lstack{C}$ is well-formed in relation to $A; \Gamma; \Delta_{N}$
        and if:
     \begin{itemize}
        \item $\lstack{C} = \cdot$ then $\feq{\Omega}{A}$.

        \item $\lstack{C} = \lframe{\Delta_a,
           p_1}{\Delta_b}{p}{\Omega_a}{\Delta_c}{\Omega_b}, \lstack{C'}$ then
           $\feq{\Omega_a}{\Omega}$ and $\Delta = \Delta_a, \Delta_b$ and
           $\Delta' = \Delta_c, p_1$

      \end{itemize}
   \end{itemize}
\end{definition}

\paragraph{Soundness of Matching} To prove the soundness of matching, we want to
reconstitute a valid match $\mz{\Gamma}{\Delta}{A}$ in HLD from the steps of
LLAM. To accomplish this, we need to prove that each step preserves 
state well-formedness.

\begin{theorem}[Transitions preserve well-formedness]
Given a rule $A \lolli B$, consider a triplet $T = A; \Gamma; \Delta_{N}$.
If a state $s_1$ is well-formed in relation to $T$ and $\trans{s_1}{s_2}$ then
$s_2$ is also well-formed.
\end{theorem}
\begin{proof}
Use case by case analysis. Straightforward use of the well-formedness
assumptions of $s_1$ and term equivalence rules.
\end{proof}

Now, we  prove the lemma that reasons about the progression of matching states.
We use the function $split(\Omega)$ that is defined as $times(flatten(\Omega))$
which de-constructs the atomic terms $A_i$ into an expression of the form $A_1
\otimes \cdots \otimes A_n$. We assume that $\feq{\Omega}{split(\Omega)}$.

\begin{lemma}[Body match result]\label{thm:body_match}
  Given a rule $A \lolli B$, consider a triplet $T = A; \Gamma; \Delta_{N}$ and
  a context $\Delta_{N} = \Delta_1, \Delta_2, \Xi$.  If $s_1 = \matstate{A
  \lolli B}{\rulestk}{\lstack{C}}{\Gamma}{\Delta_1, \Delta_2}{\Omega}{\Delta'
  \rightarrow \Omega'}$ is well-formed in relation to $T$ and
  $\transs{s_1}{s_2}$ then:

  \begin{itemize}
     \item Match succeeds with no backtracking: $s_2 = \matstate{A \lolli B}{\rulestk}{\lstack{C''},
           \lstack{C}}{\Gamma}{\Delta_1}{\cdot}{\Delta', \Delta_2 \rightarrow
              \Omega' \otimes split(\Omega)}$

     \input{lld/body}

  \end{itemize}

  If $s_1 = \contstate{A \lolli B}{\rulestk}{\lstack{C}}{\Gamma}$
  is well-formed in relation to $T$ and $\transs{s_1}{s_2}$ then either:

  \begin{itemize}
     \input{lld/body}
  \end{itemize}
\end{lemma}
\begin{proof}
   Proof by mutual lexicographic induction on the state transitions. First on the size
   of $\Omega$, then on the length of the remaining candidate facts of the
   continuation frame and then on the size of the stack $\lstack{C}$.
\end{proof}

For an initial matching state $s_1 = \matstate{A \lolli
B}{\rulestk}{\cdot}{\Gamma}{\Delta_1, \Delta_2}{A}{\cdot \rightarrow \one}$, we
know that it may transition to $s_2 = \matstate{A \lolli
B}{\rulestk}{\lstack{C'}}{\Gamma}{\Delta_1}{\cdot}{\Delta_2 \rightarrow
split(\Omega)}$. The proof $\mz{\Gamma}{\Delta_2}{split(\Omega)}$ constructed using the
steps of the LLAM is used to conclude that $\mz{\Gamma}{\Delta_2}{A}$ since
$\feq{split(\Omega)}{A}$ from the well-formedness of the machine state and by an
auxiliary theorem which proves that equivalent terms are matched in HLD using
the same facts.

\paragraph{Aggregate Soundness} Proving that an LLAM derivation is sound is
trivial except for the case of comprehensions and aggregates.  The idea is to
use lemma~\ref{thm:body_match} for the case of aggregates and use the result to
prove that the continuation stack is updated for the next iteration of the
aggregate. We sketch the proof below.

\begin{theorem}[All aggregates]\label{thm:aggregates}

Consider an aggregate $\m{agg}$, where $\Pi(\m{agg}) = \forall_{\widehat{v}, \Sigma'}.
  (\defstwo{agg}{\widehat{v}}{\Sigma'} \lolli ((\lambda x. C x)\Sigma' \with (\forall_{\widehat{x}, \sigma}.
                                               ((A \lolli B) \otimes
                                                \defstwo{agg}{\widehat{v}}{\sigma
                                                ::\Sigma'}))))$,
and a triplet $T = A; \Gamma; \Delta_{N}$.
Assume that there exists $n \geq 0$ applications of $\m{agg}$
where the $i_{th}$ application is related to the following contexts: 
  $\Delta_i$ (derived linear facts), $\Gamma_i$ (derived persistent facts),
  $\Xi_i$ (consumed linear facts),
  $V_i$ (aggregate value) and
  $\Psi_i$ (variable bindings).
The initial database context is thus $\Delta_N = \Delta, \Xi_1, \dotsc, \Xi_n$.
If $s_1 = \matstatea{\Delta, \Xi_1, \dotsc, \Xi_n}{\cdot;
  \cdot}{\Gamma}{\Delta, \Xi_1, \dotsc, \Xi_n}{A}{\cdot \rightarrow \one}{\cdot}$
  (well-formed in relation to $T$) then $\transs{s_1}{\derstatex{\Gamma}{\Delta}{\Xi, \Xi_1, \dotsc, \Xi_n}{\Gamma_{N1},                                              \Gamma_1, \dotsc, \Gamma_n}{\Delta_{N1}, \Delta_1, \dotsc,
        \Delta_n}{$\\$ (\lambda x.  C\{\Psi(\widehat{v})/\widehat{v}\} x) \Sigma,
           \Omega_N}}$ and for the $n$ aggregate applications we have:
\begin{itemize}[leftmargin=*]
     \item $n$ values $V_i$ ($\Sigma = V_n :: \dots :: V_1 :: \cdot$);
     \item $n$ soundness proofs for the $n$ aggregate matches as $\mz{\Gamma}{\Xi_i}{A}$;

     \item $n$ derivation implications for HLD that, for any $\Omega_x$ and
        $\outsem$:
  \end{itemize}
  \vspace{-2mm}
        {\small if $\dz{\Gamma}{\Pi}{\Delta, \Xi_{i+1}, \dotsc, \Xi_{n}}{\Xi, \Xi_1,
        \dotsc, \Xi_i}{\Gamma_{N1}, \Gamma_1, \dotsc, \Gamma_i}{\Delta_{N1},
           \Delta_1, \dotsc, \Delta_i}{\Omega_x}{\outsem}$ then \\
              $\dz{\Gamma}{\Pi}{\Delta, \Xi_{i+1}, \dotsc, \Xi_{n}}{\Xi, \Xi_1,
              \dotsc, \Xi_i}{\Gamma_{N1}, \Gamma_1, \dotsc, \Gamma_{i-1}}{\Delta_{N1},
              \Delta_1, \dotsc, \Delta_{i-1}}{B, \Omega_x}{\outsem}$}

\end{theorem}

\begin{proof}
Apply the corresponding body match lemma for the aggregate matching state
(similar to Lemma~\ref{thm:body_match}). If the matching process initially fails then
we have 0 aggregate applications, otherwise we have a new aggregate value $V_0$.
After updating the continuation stack, we use a generalized version of this
theorem that uses a matching state with a non-empty continuation stack
in order to retrieve the remaining conclusions for the $n-1$ aggregate applications.
\end{proof}

This theorem gives us the soundness for matching and deriving aggregates. We
use the $n$ applications of the aggregate to reconstruct the proof tree in HLD
using the rules $\dzname \m{agg}_1$, $\dzname \m{agg}_2$ and $\dzname \lolli$.
We use induction on the resulting state $s_2$ and then use the derivation
implication and the matching proof to derive the $n^{th}$ application of the
aggregate. The remaining applications are reconstructed in the same fashion
from $n-1$ to $0$ until the required HLD derivation proof is achieved.

To conclude the proof, we state the main soundness theorem that proves that a
starting LLAM state either executes a rule and builds the required HLD proof tree
or fails with no HLD proof tree.

\begin{theorem}[Soundness]\label{thm:soundness}
  If $s_1 = \dostate{\Delta_N}{\Phi}{\Gamma}{\Pi}$
  then either:
  
  \begin{itemize}
        \item $\transs{s_1}{\finalstate{\Xi}{\Gamma_{N1}}{\Delta_{N1}}}$
  and $\exists_{R \in \Phi}. \dozx{\Gamma}{\Delta_N}{\Pi}{R}{\Xi; \Gamma_{N1};
        \Delta_{N1}}$ (success)
  \item $\transs{s_1}{\failstate{\Gamma}{\Delta_N}}$ (failure)
     \end{itemize}
\end{theorem}

\begin{proof}
   Use induction on the size of the rule context $\Phi$. Apply Lemma~\ref{thm:body_match} to
   get a successful match and the required HLD proof or fail and use the
   induction hypothesis on the smaller rule context $\Phi'$.
\end{proof}
